Sound and Automated Synthesis of Digital Controllers for Physical Plants

### 2016-10-06 paper/paper.tex
hscc17

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Program Synthesis}

@@

\paragraph{General architecture of the program synthesiser}
%
The design of our synthesiser is given in
Fig.~\ref{fig:CEGIS} and consists of two phases, {\sc
Synthesise} and {\sc Verify}, which interact via a finite set of test vectors {\sc inputs}.
The {\sc synth} procedure tries to find an existential witness $P$
that satisfies the specification $\sigma(x, P)$ for all $x$:
%
$\exists P . \forall x \in \text{\sc inputs} . \sigma(x, P)$

If {\sc synthesise} succeeds in finding a witness $P$, this witness is a
candidate solution to the full synthesis formula.  We pass this candidate
solution to {\sc verify} which determines whether it does satisfy
the specification on all inputs by checking satisfiability of the
verification formula:
%
$\exists x . \lnot \sigma(x, P)$

If this formula is unsatisfiable, the candidate solution is in fact a
solution to the synthesis formula and so the algorithm terminates. 
Otherwise, the witness $x$ is an input on which the candidate solution fails
to meet the specification.  This witness $x$ is added to the {\sc inputs}
set and the loop iterates again.  It is worth noting that each iteration of the
loop adds a new input to the set of inputs being used for synthesis.  If
the full set of inputs is finite, this means that the refinement loop
can only iterate a finite number of times.

\paragraph{Elements specific to control systems}
Next, we describe the specific property that we pass to the 
program synthesiser as the specification $\sigma$.
There are two different algorithms in DSVerifier that can be used for
stability verification, one based on Schur's decomposition and another
one based on Jury's criteria \cite{DBLP:journals/dafes/BessaICF16}.
Here, we choose Jury's method due to its efficiency.

Essentially, we use this method to check the stability in the $z$-domain
for the characteristic polynomial $S(z)$ of 
the matrix $\left( \begin{array}{c} Y \\ U \end{array}\right)$.
We consider the following form for $S(z)$:
%
$$
S(z) = a_0Z^N+a_1z^{N-1}+\cdots+a_{N-1}z+a_N=0, a_0\neq0
$$
%
%\cdsay{is this for the matrix $\left( \begin{array}{c} Y \\ U \end{array}\right)$?}

Next, the following Jury matrix
$M = [m_{ij}]_{(2N−2)\times N}$ is built from $S(z)$ coefficients:
%
$$
M=\left( 
\begin{array}{c}
V^{(0)}\\
V^{(1)}\\
\vdots\\
V^{(N-2)}
\end{array}
\right)
$$
%
where $V^{(k)} = [v^{(k)}_{ij} ]_{2\times N}$ such that:
%
$$
v_{ij}^{(0)}=\left\{
\begin{array}{ll}
a_{j-1}, & \texttt{if}~i=1\\
v_{(1)(N-j+1)}^{0},&\texttt{if}~i=2
\end{array}
\right.
$$
%
$$
v_{ij}^{(k)}=\left\{
\begin{array}{ll}
0,&\texttt{if}~j>n-k\\
v_{1j}^{(k-1)}-v_{2j}^{(k-1)} . \frac{v_{11}^{(k-1)}}{v_{21}^{(k-1)}}, & \texttt{if}~j\leq n-k ~\texttt{and}~i=1\\
v_{(1)(N-j+1)}^{k},& \texttt{if}~j\leq n-k ~\texttt{and}~i=2\\
\end{array}
\right.
$$
%
where $k \in \mathbb{Z}$, such that $0 < k < N - 2$. 
$S(z)$ is the
characteristic polynomial of a stable system if and only if
the following four propositions hold:
\begin{itemize}
\item $R_1: S(1) > 0$;
\item $R_2: (−1)^N S(−1) > 0$;
\item $R_3: |a_0| < a_N$;
\item $R_4: m_{11} > 0 \wedge m_{31}>0 \wedge m_{51}>0 \wedge \cdots \wedge m_{(2N{-}3)(1)}>0$.
\end{itemize}

The stability property is then encoded by creating a
constraint of the form:
$$\phi_{stability} \equiv (R_1 \wedge R_2 \wedge R_3 \wedge R_4)$$
%\cdsay{figure out what is existentially quantified for the final $\sigma$.}

In the {\sc verify} phase, if the system is unstable for the current candidate solution,
i.e. if $\neg \phi_{stability}$ is satisfiable for $S(z)$, 
then we obtain a counterexample denoted by 
values for each coefficients of $\Delta N_{G}(z)$ and $\Delta D_{G}(z)$,
which make the closed-loop system unstable. This uncertainty is added to 
the set {\sc inputs} such that, in the subsequent {\sc synthesise} phase, 
we obtain a candidate solution consisting 
of a controller $D_C$ and $N_C$, which makes the closed-loop 
system stable for all the uncertainties accumulated in {\sc inputs}.

\paragraph{Using the program synthesiser for the running example}
We will illustrate our synthesis approach on the running example 
in Sec.~\ref{sec:running-ex}.
We start with a candidate solution that has all coefficients 0.
In the first {\sc verify} phase we obtain the following counterexample:
$$
\begin{array}{ll}
\Delta N_G(z) = 0.02634525299072265625\\
\Delta D_G(z) = 0.9979267120361328125z-0.99768829345703125
\end{array}
$$

We add this counterexample to {\sc inputs} and initiate the {\sc
  synthesise} phase, where we get the following candidate solution:

$$
\begin{array}{ll}
N_C(z) {=} 4.00888156890869140625z^2{-}0.00041961669921875z\\
\qquad\qquad -0.000213623046875\\
D_C(z) {=} -0.003753662109375z^2{+}7.37355804443359375z \\
\qquad\qquad -7.2847900390625\\
\end{array}
$$

In the second {\sc verify} phase, we check whether the current candidate
solution above is the final solution. It is not and we get a new counterexample:
$$
\begin{array}{ll}
\Delta N_G(z) = 0.02644824981689453125\\
\Delta D_G(z) {=} 1.00053310394287109375z{-}1.0014629364013671875
\end{array}
$$

After adding this new counterexample to the set of {\sc inputs}, in the 
next {\sc synthesise} phase, we get a new candidate solution that holds
for both uncertainties in {\sc Inputs}:

@@

   \path
    ([yshift=2em]synth.east) edge node {Candidate solution} ([yshift=2em]verif.west)
    ([yshift=-2em]verif.west) edge node {Counter-ex} ([yshift=-2em]synth.east)
    ([xshift=5em]verif.south) edge node {Candidate $P$} ([xshift=5em]bmc.north)
    ([xshift=-5em]bmc.north) edge node[align=center]  {UNSAT/\\model} ([xshift=-5em]verif.south)
    (verif) edge node {} (done)
    ([xshift=5em]synth.south) edge node[align=center] {Inputs} ([xshift=5em]gp.north)
    ([xshift=-5em]gp.north) edge node[align=center] {UNSAT/\\candidate} ([xshift=-5em]synth.south);
 \end{tikzpicture}
}
 \caption{Counterexample-Guided Inductive Synthesis. \label{fig:CEGIS}}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Experimental Evaluation}

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{\red{Automated} Program Synthesis for \red{for Controllers}}

@@

\paragraph{General architecture of the program synthesiser}
%
\red{[I would recommend to clarify the domain of definition of $\sigma$, particularly over the set where $x$ ranges over. 
(Be cognisant that control audience might be unfamiliar with the cegis architecture.) ]}
% 
The design of our synthesiser is given in
Fig.~\ref{fig:CEGIS} and consists of two phases, {\sc
Synthesise} and {\sc Verify}, which interact via a finite set of test vectors {\sc inputs} \blue{that is updated incrementally}.
The {\sc synth} procedure tries to find an existential witness $P$
that satisfies the specification $\sigma(x, P)$ for all $x$, namely:
%
$\exists P . \forall x \in \text{\sc inputs} . \sigma(x, P)$

If {\sc synthesise} succeeds in finding a witness $P$, this witness is a
candidate solution to the full synthesis formula.  We pass this candidate
solution to {\sc verify}, which determines whether it does satisfy
the specification on all inputs by checking satisfiability of the following 
verification formula:
%
$\exists x . \lnot \sigma(x, P)$

If this formula is unsatisfiable, the candidate solution is in fact a
solution to the synthesis formula and so the algorithm terminates. 
Otherwise, the witness $x$ is an input on which the candidate solution fails
to meet the specification.  This witness $x$ is added to the {\sc inputs}
set, and the loop iterates again.  It is worth noting that each iteration of the
loop adds a new input to the \red{finite} set $\text{\sc inputs}$  being used for synthesis.  If
the full set of inputs \blue{[which by now should have a name, as suggested to be introduced above]} is finite, this means that the refinement loop
can only iterate a finite number of times.

\paragraph{Synthesis elements specific to control systems}
Next, we describe the specific property that we pass to the 
program synthesiser as the specification $\sigma$.
There are two different algorithms in DSVerifier that can be used for
stability verification, one based on Schur's decomposition and another
one based on Jury's criteria \cite{DBLP:journals/dafes/BessaICF16} \red{[Jury here is our friend, I wouldn't use his first name ... ]}.
Here, we choose Jury's method, due to its efficiency.

We use this method to check the stability in the $z$-domain for the characteristic polynomial $S(z)$ of 
the matrix $\left( \begin{array}{c} Y \\ U \end{array}\right)$ \red{[unclear: what matrix? We need to clarify the relationship between the char. poly. $S(z)$ and the later quantities $\Delta N_{G}(z), \Delta D_{G}(z)$ and corresponding controller's structure.]}. 
We consider the following form for $S(z)$:
%
$$
S(z) = a_0Z^N+a_1z^{N-1}+\cdots+a_{N-1}z+a_N=0, a_0\neq0
$$
%
%\cdsay{is this for the matrix $\left( \begin{array}{c} Y \\ U \end{array}\right)$?}

Next, the following Jury matrix
$M = [m_{ij}]_{(2N−2)\times N}$ is built from $S(z)$ coefficients:
%
$$
M=\left( 
\begin{array}{c}
V^{(0)}\\
V^{(1)}\\
\vdots\\
V^{(N-2)}
\end{array}
\right), 
$$
%
where $V^{(k)} = [v^{(k)}_{ij} ]_{2\times N}$ such that:
%
$$
v_{ij}^{(0)}=\left\{
\begin{array}{ll}
a_{j-1}, & \texttt{if}~i=1\\
v_{(1)(N-j+1)}^{0},&\texttt{if}~i=2
\end{array}
\right.
$$
%
$$
v_{ij}^{(k)}=\left\{
\begin{array}{ll}
0,&\texttt{if}~j>n-k\\
v_{1j}^{(k-1)}-v_{2j}^{(k-1)} . \frac{v_{11}^{(k-1)}}{v_{21}^{(k-1)}}, & \texttt{if}~j\leq n-k ~\texttt{and}~i=1\\
v_{(1)(N-j+1)}^{k},& \texttt{if}~j\leq n-k ~\texttt{and}~i=2\\
\end{array}
\right.
$$
%
where $k \in \mathbb{Z}$, such that $0 < k < N - 2$. 
$S(z)$ is the
characteristic polynomial of a stable system if and only if the following four conditions hold:
\begin{itemize}
\item $R_1: S(1) > 0$;
\item $R_2: (−1)^N S(−1) > 0$;
\item $R_3: |a_0| < a_N$;
\item $R_4: m_{11} > 0 \wedge m_{31}>0 \wedge m_{51}>0 \wedge \cdots \wedge m_{(2N{-}3)(1)}>0$.
\end{itemize}

The stability property is then encoded by creating a
constraint of the form:
$$
\phi_{stability} \equiv (R_1 \wedge R_2 \wedge R_3 \wedge R_4).
$$
%\cdsay{figure out what is existentially quantified for the final $\sigma$.} 
\red{[Yes, intuitively all clear, however for the sake of clarity I would like to see the explicit mapping between the above formula and the previously used specification 
$\sigma(x, P)$, with clear elucidation of the relationship between the domain of $z$ and that of $x$.]}

In the {\sc verify} phase, if the system is unstable for the current candidate solution,
i.e. if $\neg \phi_{stability}$ is satisfiable for $S(z)$, 
then we obtain a counterexample denoted by 
values for each coefficients of $\Delta N_{G}(z)$ and $\Delta D_{G}(z)$ \red{[see my comment above - these quantities pop up too abruptly.]},
which make the closed-loop system unstable. This uncertainty is added to 
the set {\sc inputs} such that, in the subsequent {\sc synthesise} phase, 
we obtain a candidate solution consisting 
of a controller $D_C$ and $N_C$, which makes the closed-loop 
system stable for all the uncertainties accumulated in {\sc inputs}.

\paragraph{Program synthesis for the running example}
We will illustrate our synthesis approach on the running example 
in Sec.~\ref{sec:running-ex}.
We start with a candidate solution that has all coefficients 0 \red{[Be more explicit, write out spurious candidate solution with zero coefficients. KISS - you don't want to lose the control audience here.]}.
In the first {\sc verify} phase we obtain the following counterexample:
$$
\begin{array}{ll}
\Delta N_G(z) = 0.02634525299072265625\\
\Delta D_G(z) = 0.9979267120361328125z-0.99768829345703125
\end{array}
$$

We add this counterexample to {\sc inputs} and initiate the {\sc synthesise} phase, where we get the following candidate solution: 
\red{[ How is the order of the candidate solution chosen here?]}
$$
\begin{array}{ll}
N_C(z) {=} 4.00888156890869140625z^2{-}0.00041961669921875z\\
\qquad\qquad -0.000213623046875\\
D_C(z) {=} -0.003753662109375z^2{+}7.37355804443359375z \\
\qquad\qquad -7.2847900390625\\
\end{array}
$$

In the second {\sc verify} phase, we check whether the current candidate
solution above is the final solution. It is not and we get a new counterexample:
$$
\begin{array}{ll}
\Delta N_G(z) = 0.02644824981689453125\\
\Delta D_G(z) {=} 1.00053310394287109375z{-}1.0014629364013671875
\end{array}
$$
\blue{[Nice, now I'm in the groove. Next unfolds well.] }
After adding this new counterexample to the set of {\sc inputs}, in the 
next {\sc synthesise} phase, we get a new candidate solution that holds
for both uncertainties in {\sc Inputs}:

@@

   \path
    ([yshift=2em]synth.east) edge node {Candidate solution} ([yshift=2em]verif.west)
    ([yshift=-2em]verif.west) edge node {Counterexample} ([yshift=-2em]synth.east)
    ([xshift=5em]verif.south) edge node {Candidate $P$} ([xshift=5em]bmc.north)
    ([xshift=-5em]bmc.north) edge node[align=center]  {UNSAT/\\model} ([xshift=-5em]verif.south)
    (verif) edge node {} (done)
    ([xshift=5em]synth.south) edge node[align=center] {Inputs} ([xshift=5em]gp.north)
    ([xshift=-5em]gp.north) edge node[align=center] {UNSAT/\\candidate} ([xshift=-5em]synth.south);
 \end{tikzpicture}
}
 \caption{Counterexample-Guided Inductive Synthesis of Programs (CEGIS paradigm). \label{fig:CEGIS}}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Experimental Evaluation}

@@

\red{[All very interesting, however we never start from a $G(s)$ plant, we never really display the whole capabilities of the technique! Or am I missing something?]}
